\chapter*{Abstract}
\chaptermark{Abstract}
\addcontentsline{toc}{chapter}{Abstract} 

% self-contained and less than 1 pg!
% no references

% GW, motivation
Gravitational waves are ``ripples'' in spacetime emitted by massive astrophysical events. % that they carry away information about.
Over the past decade, interferometric detectors have been used to measure gravitational waves from the binary mergers of black holes and neutron stars to learn more about such systems; these gravitational waves had frequencies around 100~Hz.
Other frequencies of gravitational waves are thought to exist and contain valuable information but are yet to be detected. For example, detecting kilohertz (1--4~kHz) gravitational waves from binary neutron-star mergers could be used to further constrain the neutron-star equation-of-state and better understand exotic states of matter.
% However, gravitational waves are predicted to be emitted from other sources but are yet to be detected because they are outside the sensitive frequency range, and therefore there is valuable astrophysical information that is currently going uncollected. For example, 1--4~kHz gravitational waves from the merger and post-merger remnant of binary neutron star systems are predicted to contain information that could be used to further constrain the neutron star equation-of-state and better understand the exotic states of matter within.
% The ability to detect kilohertz (around 1~kHz) gravitational waves could also lead to better understanding of the origin of low-mass black holes, the post-bounce dynamics of core-collapse supernovae, and the discovery of exotic or primordial sources in the stochastic gravitational-wave background.
% why they can't be seen
% existing proposals for kHz sensitivity
However, to do so, the sensitivity of current detectors will need to be extended from 100~Hz to the kilohertz regime. 
% Current interferometric gravitational-wave detectors are not as sensitive at kilohertz compared to 100~Hz because of their use of optical cavities  and a limit on their quantum noise--limited sensitivity, where the quantum noise in the measurement comes from the fundamental quantum uncertainties in the state of the laser light.
% Two existing proposals to improve kilohertz sensitivity are (1) degenerate internal squeezing that uses quantum squeezing inside the detector to improve the quantum noise or (2) stable optomechanical filtering that counters the cavity resonance to improve the response to the gravitational-wave signal. However, these two proposals require low optical loss and mechanical loss, respectively.
%a configuration that combines these two existing proposals and has not been thoroughly examined to date.

The kilohertz sensitivity of current gravitational-wave detectors is limited by quantum noise from the fundamental quantum uncertainties in the state of light inside the detector. This noise can be mitigated by replacing the vacuum fluctuations entering the readout port of the detector with squeezed states.
In this thesis, I investigate a new technique to improve kilohertz sensitivity by placing a nondegenerate squeezer inside the detector. This technique, called nondegenerate internal squeezing, improves sensitivity by amplifying the detector's response to the gravitational-wave signal more than it increases the quantum noise.
To assess its feasibility, I derive an analytic Hamiltonian model of nondegenerate internal squeezing and calculate its sensitivity and stability as well as analyse its tolerance to the realistic optical losses expected in a future gravitational-wave detector.

% I investigate nondegenerate internal squeezing, a configuration that uses a nonlinear crystal to manipulate the state of the light inside the detector. This technique, called ``squeezing'', avoids the above limit on the quantum noise--limited sensitivity. Nondegenerate internal squeezing improves the quantum noise--limited sensitivity by amplifying the response to the gravitational-wave signal more than it increases the quantum noise. However, it has not been thoroughly examined to date.
% conclusions
% and show that it is robust to detection loss in the output chain. I also compare its loss tolerance to other proposed configurations to improve kilohertz sensitivity.

My model indicates that nondegenerate internal squeezing is stable, robust to detection loss in the readout, and provides a viable alternative to other proposals to improve kilohertz sensitivity. %I determine its squeezing threshold which to find the maximum 
% I validate my model by showing that it reduces to the correct high and low loss limits.
I demonstrate a technique to determine its squeezing threshold and, therefore, the limits of its operation. 
% For a realistic future gravitational-wave detector, I find that nondegenerate internal squeezing is more detection loss-resistant than degenerate internal squeezing and is a viable, all-optical alternative to stable optomechanical filtering.
I find that nondegenerate internal squeezing could feasibly improve the sensitivity of a future detector to 1--4~kHz gravitational waves. I also explore an alternative readout scheme that is promising for broadband 0.1--4~kHz sensitivity.


