\chapter*{Acknowledgements}
\chaptermark{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements} 

% supervisors, for supervision during the research component and editorial assistance
I thank my supervisor David McClelland for offering his strategy and wisdom for this project, fostering my general outlook on research, and making me feel welcome.
I thank my other supervisors (nominal and honorary) Min~Jet~Yap, Vaishali~Adya, and Sheon~Chua for their generous attention, advice, encouragement, knowledge, and humour during the research and editing of this thesis.
% squeezer research group for general discussion and support
I thank the CGA~squeezer~group for our discussions and particularly Dan Gould for his expertise with lasing theory. %mumbo jumbo
In summary, to all of the squeezer group, I quote Miranda from Shakespeare's \emph{The Tempest}:
\begin{quote}
O, wonder! \\ How many goodly creatures are there here! \\ How beauteous mankind is! O brave new world \\ That has such people in't!
\end{quote}

% collaboration: Xiang Li and Carl Bender
I thank Xiang~Li for granting the use of the code from Ref.~\cite{liBroadbandSensitivityImprovement2020} to compare to my results. I thank Carl~Bender for the stimulating discussion about PT-symmetry.
% editorial: dad
I thank Henry~Gardner and Callum~Sambridge for their editorial assistance.
% component library for pictures
The optical configurations in this thesis were illustrated using graphics from Alexander~Franzen~\cite{ComponentLibrary}.

% friends and family
Finally, I thank my friends and family for their support.


